Characterize the fetal kidney reference from the kidney cell atlas

### 2024-12-13 analyses/cell-type-ewings/template_notebooks/auc-singler-workflow/02-singler-results.Rmd

@@

```{r}
full_celltype_heatmap(classification_df, marker_gene_columns, "singler_tumor_normal")
```

```{r}
plot_density(
  classification_df,
  "tumor_sum",
  "singler_tumor_normal"
)
```

@@

```{r}
full_celltype_heatmap(classification_df, geneset_columns, "singler_tumor_normal")
```

```{r, fig.height=10}
geneset_columns |>
  purrr::map(\(column){
    plot_density(
      classification_df,
      column,
      "singler_tumor_normal"
    )
  }) |>
  patchwork::wrap_plots(ncol = 1)
```

@@

```{r}
full_celltype_heatmap(classification_df, marker_gene_columns, "singler_lumped")
```

```{r, fig.height=10}
marker_gene_columns |>
  purrr::map(\(column){
    plot_density(
      classification_df,
      column,
      "singler_lumped"
    )
  }) |>
  patchwork::wrap_plots(ncol = 1)
```

@@

```{r}
full_celltype_heatmap(classification_df, geneset_columns, "singler_lumped")
```

```{r, fig.height=10}
geneset_columns |>
  purrr::map(\(column){
    plot_density(
      classification_df,
      column,
      "singler_lumped"
    )
  }) |>
  patchwork::wrap_plots(ncol = 1)
```

### Tumor cells annotatd by `SingleR` and `AUCell`

@@

```{r}
full_celltype_heatmap(classification_df, marker_gene_columns, "consensus")
```

```{r}
plot_density(
  classification_df,
  "tumor_sum",
  "consensus"
)
```

@@

```{r}
full_celltype_heatmap(classification_df, geneset_columns, "consensus")
```

```{r, fig.height=10}
geneset_columns |>
  purrr::map(\(column){
    plot_density(
      classification_df,
      column,
      "consensus"
    )
  }) |>
  patchwork::wrap_plots(ncol = 1)
```

>>>>

```{r}
# check that marker genes are expressed, otherwise turn off those plots
total_exp <- sum(classification_df[marker_gene_columns])
if(total_exp > 0){
  show_marker_gene_plots <- TRUE
} else {
  show_marker_gene_plots <- FALSE
  message("No expression of marker genes in this library. No plots will be displayed in sections labeled 'Marker gene expression'.")
}

```

```{r}
# check that gene set scores aren't all 0, otherwise turn off those plots
total_score <- sum(classification_df[geneset_columns])
if(total_score > 0){
  show_gene_set_plots <- TRUE
} else {
  show_gene_set_plots <- FALSE
  message("Genes present in provided gene sets are not expressed in this library. No plots will be displayed in sections labeled 'Gene set scores'.")
}

```

@@

```{r, eval=show_marker_gene_plots}
full_celltype_heatmap(classification_df, marker_gene_columns, "singler_tumor_normal") 
```

```{r, eval=show_marker_gene_plots}
plot_density(
  classification_df,
  "tumor_sum",
  "singler_tumor_normal"
)
```

@@

```{r, eval=show_gene_set_plots}
full_celltype_heatmap(classification_df, geneset_columns, "singler_tumor_normal")
```

```{r, fig.height=10, eval=show_gene_set_plots}
geneset_columns |>
  purrr::map(\(column){
    plot_density(
      classification_df,
      column,
      "singler_tumor_normal"
    )
  }) |>
  patchwork::wrap_plots(ncol = 1)
```

@@

```{r, eval=show_marker_gene_plots}
full_celltype_heatmap(classification_df, marker_gene_columns, "singler_lumped")
```

```{r, fig.height=10, eval=show_marker_gene_plots}
marker_gene_columns |>
  purrr::map(\(column){
    plot_density(
      classification_df,
      column,
      "singler_lumped"
    )
  }) |>
  patchwork::wrap_plots(ncol = 1)
```

@@

```{r, eval=show_gene_set_plots}
full_celltype_heatmap(classification_df, geneset_columns, "singler_lumped")
```

```{r, fig.height=10, eval=show_gene_set_plots}
geneset_columns |>
  purrr::map(\(column){
    plot_density(
      classification_df,
      column,
      "singler_lumped"
    )
  }) |>
  patchwork::wrap_plots(ncol = 1)
```

### Tumor cells annotated by `SingleR` and `AUCell`

@@

```{r, eval=show_marker_gene_plots}
full_celltype_heatmap(classification_df, marker_gene_columns, "consensus")
```

```{r, eval=show_marker_gene_plots}
plot_density(
  classification_df,
  "tumor_sum",
  "consensus"
)
```

@@

```{r, eval=show_gene_set_plots}
full_celltype_heatmap(classification_df, geneset_columns, "consensus")
```

```{r, fig.height=10, eval=show_gene_set_plots}
geneset_columns |>
  purrr::map(\(column){
    plot_density(
      classification_df,
      column,
      "consensus"
    )
  }) |>
  patchwork::wrap_plots(ncol = 1)
```

### 2024-12-13 analyses/cell-type-ewings/template_notebooks/clustering-workflow/01-clustering-metrics.Rmd

---
title: "Clustering exploration for `r params$library_id`"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: "hide"
params:
  library_id: "SCPCL000822"
  sce_file: "../../../../data/current/SCPCP000015/SCPCS000490/SCPCL000822_processed.rds"
  cluster_results_file: "../../results/clustering/SCPCS000490/SCPCL000822_cluster-results.tsv"
---

@@

# read in results file 
cluster_results_df <- readr::read_tsv(params$cluster_results_file)
```

@@

```{r, warning=FALSE, message=FALSE}
# calculate cluster stability
stability_stats <- split_umap_df |> 
  purrr::map(\(df){ get_cluster_stability(sce, df)})
```

>>>>

---
title: "Clustering exploration for `r params$library_id`"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: "hide"
params:
  library_id: "SCPCL000822"
  sce_file: "../../../../data/current/SCPCP000015/SCPCS000490/SCPCL000822_processed.rds"
  cluster_results_file: "../../results/clustering/SCPCS000490/SCPCL000822_cluster-results.tsv"
  threads: 1
---

@@

# read in results file 
cluster_results_df <- readr::read_tsv(params$cluster_results_file) |> 
  dplyr::mutate(cluster = as.factor(cluster))
```

@@

```{r, warning=FALSE, message=FALSE}
# calculate cluster stability
stability_stats <- split_umap_df |> 
  purrr::map(\(df){ get_cluster_stability(sce, df, params$threads)})
```

### 2024-12-13 analyses/cell-type-wilms-tumor-06/notebook_template/02a_label-transfer_fetal_full_reference_Cao.Rmd

# prepare the query for label transfer
# we don't want to overwrite the srat object since `prepare_query`
#  removes features that are not present in the reference
DefaultAssay(srat) <- srat_assay
query <- prepare_query(
  srat,
  rownames(reference),
  srat_assay,
  params$homologs_file
)
```

>>>>

# prepare the query for label transfer
# we don't want to overwrite the srat object since `prepare_query`
#  removes features that are not present in the reference
DefaultAssay(srat) <- srat_assay
query <- prepare_query(
  query = srat,
  reference_rownames = rownames(reference),
  assay = srat_assay,
  homolog_file = params$homologs_file
)
```

### 2024-12-13 analyses/cell-type-wilms-tumor-06/notebook_template/03_clustering_exploration.Rmd

### Cell cycle information
```{r cell_cycle, eval=!params$testing}
s.genes <- srat@assays$RNA@meta.data$gene_ids[srat@assays$RNA@meta.data$gene_symbol %in% cc.genes$s.genes]
g2m.genes <- srat@assays$RNA@meta.data$gene_ids[srat@assays$RNA@meta.data$gene_symbol %in% cc.genes$g2m.genes]
srat <- CellCycleScoring(srat, s.features = s.genes, g2m.features = g2m.genes, set.ident = FALSE, nbin = seurat_nbins)
```

@@

We used enricher function from `clusterProfiler` to perform enrichment analysis.

>>>>

### Cell cycle information
```{r cell_cycle, eval=!params$testing}
s.genes <- srat@assays$RNA@meta.data$gene_ids[srat@assays$RNA@meta.data$gene_symbol %in% cc.genes$s.genes]
g2m.genes <- srat@assays$RNA@meta.data$gene_ids[srat@assays$RNA@meta.data$gene_symbol %in% cc.genes$g2m.genes]
# CellCycleSorting is known to throw an error when there are too few cells, e.g.:
# https://github.com/satijalab/seurat/issues/5814
# When we have fewer than ~500 cells, this error will occur. To avoid this, we override the `ctrl` (default 100) parameter to a lower value (15)
if (ncol(srat) <= 500) {
  srat <- CellCycleScoring(srat, s.features = s.genes, g2m.features = g2m.genes, set.ident = FALSE, nbin = seurat_nbins, ctrl = 15)
} else {
  srat <- CellCycleScoring(srat, s.features = s.genes, g2m.features = g2m.genes, set.ident = FALSE, nbin = seurat_nbins)
}
```

@@

We used `clusterProfiler::enricher()` to perform enrichment analysis.

### 2024-12-13 analyses/cell-type-wilms-tumor-06/supplemental_notebooks/characterize_fetal_kidney_reference_Stewart.Rmd

The input file `params$fetal_kidney_path` is the output of the script `prepare-fetal-references.R`.

## Output file

We will save the result of the differential expression analysis in `results/references/00b_marker_genes_fetal_kidney_Stewart.csv`.
Notebook is saved in the `notebook/00-reference` directory.

@@

write_csv(de_results, file = file.path(path_to_output, "00a_marker_compartment_fetal_kidney_Stewart.csv"))
```

@@

write_csv(de_results, file = file.path(path_to_output, "00a_marker_cell-type_fetal_kidney_Stewart.csv"))
```

>>>>

The input file `params$fetal_kidney_path` is the output of the script `scripts/prepare-fetal-references.R`.

## Output files

We will save the result of the differential expression analysis in `results/references/marker_<cell-type,compartment>_fetal_kidney_Stewart.csv`.

@@

write_csv(de_results, file = file.path(path_to_output, "marker_compartment_fetal_kidney_Stewart.csv"))
```

@@

write_csv(de_results, file = file.path(path_to_output, "marker_cell-type_fetal_kidney_Stewart.csv"))
```
